Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-03 paper/Abstract.tex
stuff

Correct implementation of cache coherent systems of high importance in
industrial use. While significant efforts have been made in formally
verifying that the underlying protocol, this provides no formal
guarantees to the actual physically implemented hardware cache
coherence engine, which are generally derived manually and entail
nontrivial design-oriented refinements, \eg{} buffer sizing, caching
policies, multiplexing of distinct communication channels. Previously,
it was shown that that a formal model protocol description could be
almost automatically translated into a hardware description allowing
protocol proof to be reused for the implementation; however, any
hardware-oriented change to the design resulted in a new protocol and
required a complete reverification. We propose a metholoogy where the
proof is structured to be modular and shown mechnically allowing
designers to make desired changes to the implementation without
inducing further verification overhead. We show this via the concrete
example of an hierarchical MSI protocol and many interesting design variants.

>>>>

Correct implementation of cache coherent systems is of high importance in
industrial use. Model-checking is inadequate for verifying a large system with
hierarchical caches because of its size and complexity. We provide a
theorem-prover verified proof for the safety properties of a size-parameterized
hierarchical MSI-based coherent system. The system is specified in terms of
atomic transitions accessing only local state.  Previously, it has been shown
that these atomic transitions can be translated automatically into efficient
hardware. Verifying the specification based on atomic transitions therefore
guarantees the correctness of the actual implementation.
The proof is built in a modular fashion, identifying key invariants common with
other protocols like MESI, MOSI, \etc. We informally show how these other
protocols can be proved reusing modules from the proof of the MSI protocol.

### 2014-02-03 paper/System.tex
stuff

Input and output happen at the leaves in the form of memory requests
issued by the associated processor. These will be represented by
unbounded input and output arrays. For each leaf $c$, $c.\pos$ represents the
position of the next request to process; it is set to $0$ initially.

@@

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It first involves a cache sending an upgrade request to the
its parent when the cache is not already waiting for a response for that
address from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq), and if the rest of the parent's children caches are
\emph{compatible}, and the parents state also has the permission being
requested, it sends an upgrade response to the cache, while changing the
appropriate directory state. Compatibility of an upgrade request with respect
to a cache's siblings can be defined as follows:
\begin{multline*}
\compat = \lambda (p \in Cache) (c \in Cache) (a \in \textit{Addr}) (s \in \{\Mo, \Sh, \In\}) \Rightarrow \forall i, i \ne c @-> \\
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child cache receives the response from the parent (ChildRecvResp)
and sets its state while reseting its wait state if the desired upgrade is
indicated by the response.

>>>>

Input and output happen at the leaves in the form of memory requests
issued by and responses issued to the associated processor. These will be
represented by two mappings one from $\mathbb{N}$ to requests and the other
from $\mathbb{N}$ to responses. For each leaf $c$, $c.\pos$ represents the
position of the next request to process; it is set to $0$ initially.

@@

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It first involves a cache sending an upgrade request to the
its parent when the cache is not already waiting for a response for that
address from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq), and if the rest of the parent's children caches are
\emph{compatible}, and the parents state also has the permission being
requested, it sends an upgrade response to the cache, while changing the
appropriate directory state. Compatibility of an upgrade request with respect
to a cache's siblings can be defined as follows:
\begin{multline*}
\compat = \lambda (p \in Cache) (c \in Cache) (a \in \textit{Addr}) (s \in
\{\Mo, \Sh, \In\}) \Rightarrow\\ \forall i, i \ne c @->
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child cache receives the response from the parent (ChildRecvResp)
and sets its state while reseting its wait state if the desired upgrade is
indicated by the response.

### 2014-02-03 paper/Title.tex
stuff

\title{Proof Modularization: Enabling Verified Design-Oriented Modifications in Cache Coherence Implementation}

@@

%\centerline{\footnotesize {\bf Keywords}: Guarded Atomic Actions, HW Synthesis, Modular Composition}

>>>>

\title{Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof}

% hierarchical cache coherence
% modular proof

@@

\centerline{\footnotesize {\bf Keywords}: Cache coherence, theorem prover, cache hierarchy, guarded atomic actions}
